A thesis draft

### 2018-11-20 res/sections/08-Background.tex
Add part on encapsulation and continue part on SFC

\section{Service Function Chaining}
The Service Function Chaining (SFC) concept was defined  in~\cite{halpern2015service} and refers to an ordered set of Service Functions (SF) and the constraint, defined after a classification process, to packets and flows. SF are the particular functions whose aim is the treatment of packets received. They can work on different levels of the network layer. These components can be implemented as virtual element or a combination of a physical network element and the software running on the top of it. There are no limitations on the number of SFs running on the same network element, so in general it is possible to have one or more functions running on the same element. SFs can be aware or not that are part of a certain chain, in the matter case they are called \emph{SFC aware} functions, in the latter \emph{SFC unaware}.

An SFC can be seen as a graph in which each node is an SF and edges are the path through which a packet have to pass through. Because of this definitions there are no limitation on possible cycles that can occur in the chain, in case that a certain SF have to be reapplied on the same packet, frame or flow. These scenarios must be adequately managed to avoid endless loops. This enable even the possibility of branching on a certain node in order to define the treatment that has to be imposed from that point of the chain. The graph representation also permit to think to \emph{unidirectional} and \emph{bidirectional} chains. An SFC is defined unidirectional if traffic can traverse it only in one direction, so if it is composed by 3 SF, named \texttt{SF0}, \texttt{SF1} and \texttt{SF2} in this order, traffic can only go from \texttt{SF0} to \texttt{SF2}. On the contrary, a bidirectional SFC allows traffic in both directions. So, having the same chain as in the previous example, traffic can pass both from \texttt{SF0} to \texttt{SF2} and from \texttt{SF2} to \texttt{SF0}.

>>>>

\section{Service Function Chaining}
The Service Function Chaining (SFC) concept was defined 
in~\cite{halpern2015service} and refers to an ordered set of Service Functions
(SF) and the constraint, defined after a classification process, to packets and
flows. SF are the particular functions whose aim is the treatment of packets
received. They can work on different levels of the network layer. These
components can be implemented as virtual element or a combination of a physical
network element and the software running on the top of it. There are no
limitations on the number of SFs running on the same network element, so in
general it is possible to have one or more functions running on the same
element. SFs can be aware or not that are part of a certain chain, in the
matter case they are called \emph{SFC aware} functions, in the latter one \emph
SFC unaware}.

\subsection{Architecture principles}
The SFC technology was though to pursue some concept. One of the most important
principles is the \emph{decoupling}. Decoupling in terms on independence on the
infrastructure and the network, so no change must be performed to the
forwarding underlying topology to deploy SFCs, and in terms of separation
between packet handling (by SFs) and forwarding. This decoupling refers even to
architecture and functions: one must be independent from the other. Another
main point is the possibility to dynamically set routes on the incoming traffic
to decide the best treatment based on its typology. The classification can
occur on different point of the chain and can be updated during the packet
delivery. Finally creation, modification and deletion of a certain SFC must has
To impact on the others.

\subsection{Graph view of an SFC}
An SFC can be seen as a graph in which each node is an SF and edges are the
path through which a packet have to pass through. Because of this definitions
there are no limitation on possible cycles that can occur in the chain, in case
that a certain SF have to be reapplied on the same packet, frame or flow. These
scenarios must be adequately managed to avoid endless loops. This enable even
the possibility of branching on a certain node in order to define the treatment
that has to be imposed from that point of the chain. Traffic path in the SFC
graph are managed and described by the \emph{Service Function Forwarder} (SFF)
and \emph{Service Function Path} (SFP) components. The matter is the element of
the architecture responsible of traffic forwarding. The main goals are to
redirect traffic according to the SFC specified for a specific packet flow,
query the classifier if needed and manage traffic coming back from a SF. The
SFP, instead, is an indirection layer between the fully abstract notion of SFC
and the list of SF that the traffic has to traverse. It is used for dictate the
constraint specification on the path of where packets assigned to a certain
service function path must go. There are no limitation on how specific it must
be: it can specify the exact locations, identifying both the SFF and the SF, or
only give some recommendation or, for example, defining only the SFF and
demanding to it the decision on the SF instance.

The graph representation also permit to think to \emph{unidirectional} and
\emph{bidirectional} chains. An SFC is defined unidirectional if traffic can
traverse it only in one direction, so if it is composed by 3 SF, named
\texttt{SF0}, \texttt{SF1} and \texttt{SF2} in this order, traffic can only go
from \texttt{SF0} to \texttt{SF2}. On the contrary, a bidirectional SFC allows
traffic in both directions. So, having the same chain as in the previous
example, traffic can pass both from \texttt{SF0} to \texttt{SF2} and from
\texttt{SF2} to \texttt{SF0}.

\subsection{SFC Encapsulation}
In order to exchange packet and information in the SFC ecosystem, traffic is
encapsulated. The so called SFC encapsulation carries data used by the framework
to identify SFPs, but it differs from a transport encapsulation and it is
independent from the transport one: as such any transport protocol can be
exploited to forward traffic.

\todo{add sfc proxy}

The \emph{Network Service Header} (NSH), used by SFC environment for
encapsulation is described in~\cite{rfc8300} and consider three main fields as
represent in Figure~\ref{chap:background:img:nshformat}: Base Header, Service
Path Header and Context Header.
\begin{figure}[H]
  \centering
  \includegraphics[scale=0.6]{nshformat}
  \caption{Network Service Header~\cite{rfc8300}}
  \label{chap:background:img:nshformat}
\end{figure}

\subsubsection{Base header}
The Base Header utility is provide information about the service header and the
payload protocol.

\begin{figure}[H]
  \centering
  \includegraphics[scale=0.6]{nshbase}
  \caption{Network Service Header~\cite{rfc8300}}
  \label{chap:background:img:nshformat}
\end{figure}

Fields are:
\begin{description}
  \item[Version]: This field is used to set the version of the header, used for
  ensuring backward compatibility through different NSH definitions;
  \item[0 bit]: Bit set only in case of orchestration or management packets, and
  must not be changed by the SFC;
  \item[TTL]: Time To Live, indicate the maximum number of hops among SFF and
  its initial value should be configurable by the control plane. At each hop it
  must be decremented to 1 until 0: in the latter case the packet must be
  discarded. The default value is 63. This field is useful in loop prevention
  mechanism. The default value is 63;
  \item[Length]: the total length of the header (in 4-byte words), including
  the \texttt{Fixed-Length Context Header} in NSH type 0x1 or the
  \texttt{Variable-Length Context Header} in NSH type 0x2. In case of variable
  length header data must be padded to a multiple of 4;
  \item[Unassigned bits]: unassigned bits for future usages;
  \item[MD]: Metadata Type, indicate the type of the NSH. Values that now
  accepted are \texttt{0x0} that is a reserved value and if set the packet must
  be discarded, \texttt{0x1} indicate that context header length is
  fixed~\ref{subsubsec:background:flch}, \texttt{0x2} indicate that context
  header length is variable~\ref{subsubsec:background:vlch} and \texttt{0xF} for
  testing purposes (if not part of the experiment the packet must be discarded).
  Base header length and Service Path Header have always the same size;
  \item[Next Protocol]: indicates the protocol type of encapsulated data
  (protocols codes can be found at~\cite{rfc8300}).
\end{description}

\subsubsection{Service Path Header}
Provides path identification and location within a service path.

\begin{figure}[H]
  \centering
  \includegraphics[scale=0.6]{nshservicepath}
  \caption{NSH Service Path Header~\cite{rfc8300}}
  \label{chap:background:img:nshpath}
\end{figure}

\begin{description}
  \item[Service Path Identifier (SPI)]: 24 bits, identify unambiguously an SFP.
  The first classifier have to set it;
  \item[Service Index (SI)]: 8 bits, provide location within the SFP. Initially
  must be set to 255 and it must be decremented on each hop.
\end{description}

SPI combined to SI determine the next SF on each link in the chain. SI change
provide a minimal loop prevention mechanism.

\subsubsection{NSH with Fixed-Length Context Header}
\label{subsubsec:background:flch}
The NSH of type \texttt{0x1} have a 16-Byte Fixed-Length Context Header that
can be used to share data along the service path.

\begin{figure}[H]
  \centering
  \includegraphics[scale=0.6]{nshtype1}
  \caption{NSH MD Type \texttt{0x1}~\cite{rfc8300}}
  \label{chap:background:img:nshtype1}
\end{figure}

The specification does not specify the content of this field and if SF or SFC
Proxy does not known its semantics they have to discard the entire packet.

\subsubsection{NSH with Variable-Length Context Header}
\label{subsubsec:background:vlch}
In NSH of type \texttt{0x2}, zero or more Variable-Length Context Headers may
be added, but lengths must always be a multiple of 4-bytes (otherwise they must
be padded).

\begin{figure}[H]
  \centering
  \includegraphics[scale=0.6]{nshtype2}
  \caption{NSH MD Type \texttt{0x2}~\cite{rfc8300}}
  \label{chap:background:img:nshtype2}
\end{figure}

The structure of the variable length part is depicted in
Figure~\ref{chap:background:img:nshtype2variable}

\begin{figure}[H]
  \centering
  \includegraphics[scale=0.6]{nshtype2variable}
  \caption{Variable-Length Context Header~\cite{rfc8300}}
  \label{chap:background:img:nshtype2variable}
\end{figure}

\begin{description}
  \item[Metadata Class (MD Class)]: field used in order to provide hierarchical
  namespaces for standards, vendors or other scopes;
  \item[Type]: indicates the type of metadata, depending on the MD Class;
\end{description}

\subsection{Classification of traffic}
In order to define the path of which traffic have to be forwarded, it must be
classified. The classification granularity is not specified and can occur even
during the SFC traverse. At the ingress of SFC environment packets have to be
classified to determine to choose an SFC and eventually, if the classification
granularity is too coarse, can be repeated after the treatment of some SFs. The
reclassification, also called branching, may result in a different SFP choice or
only to a metadata change.

\subsection{Considerations}
